---
title: "lab_6"
author: "Cristina"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank Based test example 
Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test)
Comparison of ranks between two groups! 

H0: Equal ranks (or medians)
HA: These are drawn from population where ranks are not equal or medians are not equal 

### Create our mock samples 

set.seed() to create pseudorandom but reproducible samples 

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)
gp_1

set.seed(1424)

gp_2 <- sample.int(30, size =15, replace =TRUE)
gp_2
```
Take a quick look at the distribution using hist()

```{r}
hist(gp_1)
hist(gp_2)
```


Try a t-test to compare the results: 

```{r}
ttest <- t.test(gp_1, gp_2)

ttest

#If these data were paired, I can just add the argument 'paired = TRUE' within the t.test() function.
```



The p-value means: this p-value means that if our samples are drawn from populations with the same mean, (if the H0 is true) there is a `r round(ttest$p.value, 2)` probability of finding sample means *at least as different as mine* by random chance. 


A more comprehensive explanation of difference between groups...actually means, data spread, effect size.. there is no significant difference (t(df) = t.statistic value, p = x). 


#### Mann Whitney U rank based test 

We will use the wilcox.test()function to perform a two-sample rank-based test. 

```{r}
mwu <- wilcox.test(gp_1, gp_2)
mwu
```
This p-value means that if our samples are drawn from population with the ame median. there is a 28% chance of finding a sample medians at least as different as ours by random chance. 

No significant difference in medians (Mann-Whitney U; `r mwu$statistic`, p = `r round(mwu$p.value, 2)`) 

The medians of group a (M=x) and group b (M= x) did not differ significantly (Mean WHitney U....) 

## Simple linear regression 

Two continuous quantitative variables that you are trying to understand the relationship between, and you think that a linear relationship makes sense. 

Use the pengins data from palmer penguins, exploring flipper length (x) vs body mass (y). 

```{r}
# use view(penguins) to look at dataset
# Make an exploratory scatterplot of peguin flipper length (x) vs body mass (y)

ggplot(data = penguins, aes(x=flipper_length_mm, y = body_mass_g))+
  geom_point() +
  geom_smooth(method = "lm") # ADD A TRENDLINE 

# looking at the graph, does it makes sense to use a linear regression? does it look like it could follow a linear pattern? 


```
### Find the model estimates for coefficients 

```{r}
# Use the 'lm()' for linear model. 
#using model notation

penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
```

model coefficients: `r penguin_lm$coefficients`
flipper length coefficient: `r penguin_lm$coefficients[2]`


### broom tidy to get a tidy outpur from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy
```

### Use broom:: glance() to get output of model fit & statistical -- instead of working with lists, you can work with a table 

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_tidy

penguin_lm_tidy$estimate[2]
```
### Use equatiomatic:: extract_eq() to return the model in LaTex formula notation. 

```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)  # include actual coefficient values
```

